---
title: "Modeling C3 Photosynthesis: recommendations for common scenarios"
author: "Chris Muir"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Modeling C3 Photosynthesis: recommendations for common scenarios}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
biblography: 
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(photosynthesis)
```

# 'Typical' hypostomatous leaf

Here I show how to implement a model close to the most widely used ones, in which boundary layer conductances are high, cuticular conductance is 0, all stomatal conductance is through the lower (abaxial) leaf surface, and there is a single mesophyll conductance. The advantage of this approach are that is its simplicity. You can ignore several complexities such as:

* Low boundary layer conductance that varies between surfaces (can happen in large leaves at low wind speed)
* Large cuticular conductance (this can be happen in young leaves)
* Stomatal conductance can occur through both surface in amphistomatous leaves (this is common in herbs, for example)
* Mesophyll conductance can be partitioned into air and liquid phases

If these complexities are important to you, consider more complex models you can implement with **photosynthesis**. But, if you don't care, here's how to make things simple:

```{r, simple leaf, eval=FALSE}
library(photosynthesis)
bake_par   = make_bakepar()                       # temperature response parameters
constants  = make_constants(use_tealeaves = FALSE) # physical constants
# leaf parameters
leaf_par   = make_leafpar(
  replace = list(
    # Set cuticular conductance to 0
    g_uc = set_units(0, umol / (m^2 * s * Pa)),
    # All conductance through lower stomata and mesophyll
    k_sc = set_units(0),
    k_sc = set_units(0)
  ),
  use_tealeaves = FALSE
  )   
enviro_par = make_enviropar(use_tealeaves = FALSE) # environmental parameters

```


# Amphistomatous leaf

Most leaves, especially on woody plants, are hypostomatous (Muir 2015) meaning that all the stomatal conductance

# References

Muir CD. 2015. Making pore choices: repeated regime shifts in stomatal ratio. Proc Roy Soc B 282: 20151498.
