---
title: "photosynthesis-curve-fitting-sensitivity-analyses"
author: "Joseph R. Stinziano"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
vignette: >
 %\VignetteIndexEntry{photosynthesis-curve-fitting-sensitivity-analyses}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
---

This vignette is designed to demonstrate how to use the curve fitting and sensitivity analysis tools
Sections are named based on the set of methods to be used:

1. Fitting light response curves

2. Fitting CO2 response curves

3. Fitting temperature response curves (Need data & to complete tutorial here)

4. Fitting stomatal conductance models

5. Fitting light respiration

6. Fitting mesophyll conductance

7. Fitting pressure-volume curves

8. Fitting hydraulic vulnerability curves

9. Sensitivity analyses (just need to think about measures of sensitivity & multi fits)

10. Dependency checking

Components under construction:

1. Full Gu-type CO2 response fitting

2. alphag fitting

3. Busch et al (2018) CO2 response model

Within each section, data will either be generated or used from an
installed dataset within the package. For help with a given function,
please consult the help file via: ?functionname in the console. If you
want to know the fine details of the code, please go to:

https://github.com/jstinzi/photosynthesis

And look in the R folder to find the raw function files. These contain
heavily annotated code that explains the why and how of their operation.

#Installing the package

You will need the following packages:

devtools - lets you install packages from Github and Bitbucket

minpack.lm - useful for nonlinear curve fitting that is more robust
than base R

tidyverse - set of tools for manipulating data within R

FOR WINDOWS USERS

You will need to install Rtools, available at:

https://cran.r-project.org/bin/windows/Rtools/

```{r}
#To install, run the following without comments
#library(devtools)
#install_github("jstinzi/photosynthesis")

#Load package
library(photosynthesis)

#To cite, use:
citation("photosynthesis")

#Load tidyr - needed for vignette manipulations
library(tidyr)

```

#Reading Li-Cor data

If you are trying to read in the raw data files of the Li-Cor 6400 or 6800
models, you can use the package RLicor by Erik Erhardt available on Github.
```{r}
#library(devtools)
#install_github("erikerhardt/RLicor")
#library(RLicor)

#The following will detect and read Li-Cor 6400 and 6800 files
#?read_Licor

#To cite, use:
#citation("RLicor")
```

#1. Fitting light response curves

This package currently only implements the Marshall et al. 1980
non-rectangular hyperbola model of the photosynthetic light response.
```{r}
#Read in your data
#Note that this data is coming from data supplied by the package
#hence the complicated argument in read.csv()
#This dataset is a CO2 by light response curve for a single sunflower
#Note that to read in your own data, you will need to delete the
#system.file() function, otherwise you will get an error
data <- read.csv(system.file("extdata", "A_Ci_Q_data_1.csv",
                             package = "photosynthesis"))

#Fit many AQ curves
#Set your grouping variable
#Here we are grouping by CO2_s and individual
data$C_s <-(round(data$CO2_s, digits = 0))

#For this example we need to round sequentially due to CO2_s setpoints
data$C_s <- as.factor(round(data$C_s, digits = -1))

#To fit one AQ curve
fit <- fit_aq_response(data[data$C_s == 600,],
                       varnames = list(A_net = "A",
                                         PPFD = "Qin",
                                       Q_cut = 250))

#Print model summary
summary(fit[[1]])

#Print fitted parameters
fit[[2]]

#Print graph
fit[[3]]

#Fit many curves
fits <- fit_many(data = data,
                 varnames = list(A_net = "A",
                                         PPFD = "Qin",
                                         group = "C_s"),
                 funct = fit_aq_response,
                 group = "C_s")

#Look at model summary for a given fit
#First set of double parentheses selects an individual group value
#Second set selects an element of the sublist
summary(fits[[3]][[1]])

#Print the parameters
fits[[2]][[2]]

#Print the graph
fits[[3]][[3]]

#Compile graphs into a list for plotting
fits_graphs <- compile_data(fits,
                            list_element = 3)

#Print graphs to jpeg
#print_graphs(data = fits_graphs,
#             path = tempdir(),
#             output_type = "jpeg")

#Compile parameters into dataframe for analysis
fits_pars <- compile_data(fits,
                          output_type = "dataframe",
                          list_element = 2)

```

#2. Fitting CO2 response curves

This package currently implements a Gu-type fitting procedure for
CO2 response curves similar to the Duursma (2015) implementation.
There is ongoing work to implement a full Gu-type method whereby
mesophyll conductance, Km, and GammaStar could all be fit (Gu et al
2010). There is also ongoing work to implement a procedure to fit 
alphag for the TPU-limited region and to incorporate the Sharkey
(2019) suggestion of using chlorophyll fluorescence data to inform
TPU limitations.
```{r}
#Read in your data
#Note that this data is coming from data supplied by the package
#hence the complicated argument in read.csv()
#This dataset is a CO2 by light response curve for a single sunflower
data <- read.csv(system.file("extdata", "A_Ci_Q_data_1.csv", 
                             package = "photosynthesis"))

#Define a grouping factor based on light intensity to split the ACi
#curves
data$Q_2 <- as.factor((round(data$Qin, digits = 0)))

#Convert data temperature to K
data$T_leaf <- data$Tleaf + 273.15

#Fit ACi curve. Note that we are subsetting the dataframe
#here to fit for a single value of Q_2
fit <- fit_aci_response(data[data$Q_2 == 1500, ],
                        varnames = list(A_net = "A",
                                      T_leaf = "T_leaf",
                                      C_i = "Ci",
                                      PPFD = "Qin"))

#View fitted parameters
fit[[1]]

#View graph
fit[[2]]

#View data with modeled parameters attached
#fit[[3]]

#Fit many curves
fits <- fit_many(data = data,
                 varnames = list(A_net = "A",
                                      T_leaf = "T_leaf",
                                      C_i = "Ci",
                                      PPFD = "Qin"),
                 funct = fit_aci_response,
                 group = "Q_2")

#Print the parameters
#First set of double parentheses selects an individual group value
#Second set selects an element of the sublist
fits[[3]][[1]]

#Print the graph
fits[[3]][[2]]

#Compile graphs into a list for plotting
fits_graphs <- compile_data(fits,
                            list_element = 2)

#Print graphs to pdf.
#print_graphs(data = fits_graphs,
#             path = tempdir(),
#             output_type = "pdf",
#             pdf_filename = "mygraphs.pdf")

#Compile parameters into dataframe for analysis
fits_pars <- compile_data(fits,
                          output_type = "dataframe",
                          list_element = 1)

```

#3. Fitting temperature response curves

This package provides support for multiple temperature response
functions (Arrhenius 1915; Medlyn et al. 2002; Kruse & Adams. 2006;
Heskel et al. 2016; Liang et al. 2018).
```{r}
#Read in data
data <- read.csv(system.file("extdata", "A_Ci_T_data.csv", 
                             package = "photosynthesis"),
                 stringsAsFactors = FALSE)

#Round temperatures to group them appropriately
#Use sequential rounding
data$T2 <- round(data$Tleaf, 1)
data$T2 <- round(data$Tleaf, 0)

#Look at unique values to detect rounding issues
unique(data$T2)

#Some still did not round correctly,
#manually correct
for(i in 1:nrow(data)){
  if(data$T2[i] == 18){
    data$T2[i] <- 17
  }
  if(data$T2[i] == 23){
    data$T2[i] <- 22
  }
  if(data$T2[i] == 28){
    data$T2[i] <- 27
  }
  if(data$T2[i] == 33){
    data$T2[i] <- 32
  }
  if(data$T2[i] == 38){
    data$T2[i] <- 37
  }
}

#Make sure it is a character string for grouping
data$T2 <- as.character(data$T2)

#Create grouping variable by ID and measurement temperature
data <- unite(data, col = "ID2", c("ID", "T2"),
                 sep = "_")

#Split by temperature group
data <- split(data, data$ID2)

#Obtain mean temperature for group so temperature
#response fitting is acceptable later, round to
#2 decimal places
for(i in 1:length(data)){
  data[[i]]$Curve_Tleaf <- round(mean(data[[i]]$Tleaf), 2)
}

#Convert from list back to dataframe
data <- do.call("rbind", data)

#Parse grouping variable by ID and measurement temperature
data <- separate(data, col = "ID2", into = c("ID", "T2"),
                 sep = "_")

#Make sure number of values matches number of measurement
#temperatures. May vary slightly if plants had slightly
#different leaf temperatures during the measurements
unique(data$Curve_Tleaf)

#Create ID column to curve fit by ID and temperature
data <- unite(data, col = "ID2", c("ID", "Curve_Tleaf"),
                 sep = "_")

#Convert data temperature to K
data$T_leaf <- data$Tleaf + 273.15

#Fit many CO2 response curves
fits2 <- fit_many(data = data,
                 group = "ID2",
                 varnames = list(A_net = "A",
                                 C_i = "Ci",
                                 T_leaf = "T_leaf",
                                 PPFD = "Qin",
                                 g_mc = "g_mc"),
                 funct = fit_aci_response,
                 alphag = 0)

#Extract ACi parameters
pars <- compile_data(fits2, output_type = "dataframe",
                     list_element = 1)

#Extract ACi graphs
graphs <- compile_data(fits2, output_type = "list",
                     list_element = 2)

#Parse the ID variable
pars <- separate(pars, col = "ID", into = c("ID", "Curve_Tleaf"), sep = "_")

#Make sure curve leaf temperature is numeric
pars$Curve_Tleaf <- as.numeric(pars$Curve_Tleaf)
pars$T_leaf <- pars$Curve_Tleaf + 273.15
out <- fit_t_response(data = pars[pars$ID == "S2",],
                      varnames = list(Par = "V_cmax",
                                      T_leaf = "T_leaf"),
                      setvar = "Hd")

out[["Arrhenius"]][["Graph"]]
out[["Heskel"]][["Graph"]]
out[["Kruse"]][["Graph"]]
out[["Medlyn"]][["Graph"]]
out[["MMRT"]][["Graph"]]
out[["Quadratic"]][["Graph"]]
out[["Topt"]][["Graph"]]
```

#4. Fitting stomatal conductance models

The package currently supports three varieties of stomatal conductance
models (Ball et al. 1987; Leuning 1995; Medlyn et al. 2011).
```{r}
#Read in your data
#Note that this data is coming from data supplied by the package
#hence the complicated argument in read.csv()
#This dataset is a CO2 by light response curve for a single sunflower
data <- read.csv(system.file("extdata", "A_Ci_Q_data_1.csv", 
                             package = "photosynthesis"))

#Convert RH to a proportion
data$RH <- data$RHcham / 100

#Fit stomatal conductance models
#Can specify a single model, or all as below
fits <- fit_gs_model(data = data,
                     varnames = list(A_net = "A",
                                     C_air = "Ca",
                                     g_sw = "gsw",
                                     RH = "RH",
                                     VPD = "VPDleaf"),
                     model = c("BallBerry",
                               "Leuning",
                               "Medlyn_partial",
                               "Medlyn_full"),
                         D0 = 3)

#Look at BallBerry model summary:
summary(fits[["BallBerry"]][["Model"]])

#Look at BallBerry parameters
fits[["BallBerry"]][["Parameters"]]

#Look at BallBerry plot
fits[["BallBerry"]][["Graph"]]


#Fit many g_sw models
#Set your grouping variable
#Here we are grouping by Qin and individual
data$Q_2 <- as.factor((round(data$Qin, digits = 0)))

fits <- fit_many(data,
                 varnames = list(A_net = "A",
                                 C_air = "Ca",
                                 g_sw = "gsw",
                                 RH = "RH",
                                 VPD = "VPDleaf"),
                 funct = fit_gs_model,
                 group = "Q_2")

#Look at the Medlyn_partial outputs at 750 PAR
#Model summary
summary(fits[["750"]][["Medlyn_partial"]][["Model"]])

#Model parameters
fits[["750"]][["Medlyn_partial"]][["Parameters"]]

#Graph
fits[["750"]][["Medlyn_partial"]][["Graph"]]

#Compile parameter outputs for BallBerry model
#Note that it's the first element for each PAR value
#First compile list of BallBerry fits
bbmods <- compile_data(data = fits,
                     output_type = "list",
                     list_element = 1)
#Now compile the parameters (2nd element) into a dataframe
bbpars <- compile_data(data = bbmods,
                     output_type = "dataframe",
                     list_element = 2)

#Convert group variable back to numeric
bbpars$ID <- as.numeric(bbpars$ID)

#Take quick look at light response of intercept parameters
plot(g0 ~ ID, bbpars)

#Compile graphs
graphs <- compile_data(data = bbmods,
                       output_type = "list",
                       list_element = 3)

#Look at 3rd graph
graphs[[3]]
```

#5. Fitting light respiration

The package currently supports fitting light respiration according to
the slope-intercept regression extension of the Laisk method (Laisk 
1977; Walker & Ort 2015), the Kok method (Kok 1956), and the Yin
modification of the Kok method (Yin et al. 2009, 2011)
```{r}
#Read in your data
#Note that this data is coming from data supplied by the package
#hence the complicated argument in read.csv()
#This dataset is a CO2 by light response curve for a single sunflower
data <- read.csv(system.file("extdata", "A_Ci_Q_data_1.csv", 
                             package = "photosynthesis"))

#Fit light respiration with Yin method
r_light <- fit_r_light_yin(data = data,
                           varnames = list(A_net = "A",
                                           PPFD = "Qin",
                                           phi_PSII = "PhiPS2"),
                           PPFD_lower = 20,
                           PPFD_upper = 250)

#Fit light respiration with Kok method
r_light <- fit_r_light_kok(data = data,
                           varnames = list(A_net = "A",
                                           PPFD = "Qin"),
                           PPFD_lower = 20,
                           PPFD_upper = 150)

#Set your grouping variable
#Here we are grouping by CO2_s and individual
data$C_s <-(round(data$CO2_s, digits = 0))

#For this example we need to round sequentially due to CO2_s setpoints
data$C_s <- as.factor(round(data$C_s, digits = -1))

#Fit light respiration across groups with Yin method
r_lights <- fit_many(data = data,
                     funct = fit_r_light_yin,
                     group = "C_s",
                     varnames = list(A_net = "A",
                                           PPFD = "Qin",
                                           phi_PSII = "PhiPS2"),
                           PPFD_lower = 20,
                           PPFD_upper = 250)

#Compile the outputs - note this is slightly more complex because the
#output of the fit_many above is a list of atomic vectors, not dataframes.
group <- names(r_lights)
r_lights <- do.call("c", r_lights)
r_light_yin <- data.frame(x = group, y = r_lights, stringsAsFactors = FALSE)
r_light_yin$x <- as.numeric(r_light_yin$x)
colnames(r_light_yin) <- c("C_s", "r_light")

#Fit the Walker-Ort method for GammaStar and light respiration
walker_ort <- fit_r_light_WalkerOrt(data,
                      varnames = list(A_net = "A",
                                      C_i = "Ci",
                                      PPFD = "Qin"))

#View model output
summary(walker_ort[[1]])

#View graph
walker_ort[[2]]

#View coeffients
walker_ort[[3]]

```

#6. Fitting mesophyll conductance

Currently there is only support for fitting mesophyll conductance
according to the variable J method from Harley et al. 1992
```{r}
#Read in your data
#Note that this data is coming from data supplied by the package
#hence the complicated argument in read.csv()
#This dataset is a CO2 by light response curve for a single sunflower
data <- read.csv(system.file("extdata", "A_Ci_Q_data_1.csv", 
                             package = "photosynthesis"))

#Note: there will be issues here if the alpha value used
#for calculating ETR is off, if GammaStar is incorrect,
#if Rd is incorrect.
data <- fit_g_mc_variableJ(data,
                         varnames = list(A_net = "A",
                                         J_etr = "ETR",
                                         C_i = "Ci",
                                         PPFD = "Qin",
                                         phi_PSII = "PhiPS2"),
                                         gamma_star = 46,
                                         R_d = 0.153,
                         usealpha_Q = TRUE,
                         alpha_Q = 0.84,
                         beta_Q = 0.5,
                         P = 84)

#Note that many g_mc values from this method can be unreliable
ggplot(data, aes(x = CO2_s, y = g_mc, colour = reliable)) +
  labs(x = expression(CO[2]~"("*mu*mol~mol^{-1}*")"),
       y = expression(g[m]~"(mol"~m^{-2}~s^{-1}~Pa^{-1}*")")) +
  geom_point(size = 2) +
  theme_bw() +
  theme(legend.position = 'bottom')

#Plot QAQC graph according to Harley et al. 1992
ggplot(data, aes(x = CO2_s, y = dCcdA, colour = reliable)) +
  labs(x = expression(CO[2]~"("*mu*mol~mol^{-1}*")"),
       y = expression(delta*C[c]*"/"*delta*A)) +
  geom_hline(yintercept = 10) +
  geom_point(size = 2) +
  theme_bw() +
  theme(legend.position = 'bottom')

ggplot(data, aes(x = dCcdA, y = g_mc, colour = reliable)) +
  labs(x = expression(delta*C[c]*"/"*delta*A),
       y = expression(g[m]~"(mol"~m^{-2}~s^{-1}~Pa^{-1}*")")) +
  geom_point(size = 2) +
  theme_bw() +
  theme(legend.position = 'bottom')
```

#7. Fitting pressure-volume curves

This package follows the Prometheus wiki spreadsheet from Sack and Pasquet-Kok at:

http://prometheuswiki.org/tiki-index.php?page=Leaf+pressure-volume+curve+parameters. 

For references, see Koide et al. 2000, Sack et al. 2003, and Tyree & Hammel 1972.
```{r}
#Read in data
data <- read.csv(system.file("extdata", "PV_curve.csv", 
                             package = "photosynthesis"))

#Fit one PV curve
fit <- fit_PV_curve(data[data$ID == "L2", ],
                    varnames = list(psi = "psi", 
                                    mass = "mass", 
                                    leaf_mass = "leaf_mass", 
                                    bag_mass = "bag_mass", 
                                    leaf_area = "leaf_area"))

#See fitted parameters
fit[[1]]

#Plot water mass graph
fit[[2]]

#Plot PV Curve
fit[[3]]

#Fit all PV curves in a file
fits <- fit_many(data,
                 group = "ID",
                 funct = fit_PV_curve,
                 varnames = list(psi = "psi", 
                                    mass = "mass", 
                                    leaf_mass = "leaf_mass", 
                                    bag_mass = "bag_mass", 
                                    leaf_area = "leaf_area"))

#See parameters
fits[[1]][[1]]

#See water mass - water potential graph
fits[[1]][[2]]

#See PV curve
fits[[1]][[3]]

#Compile parameter outputs
pars <- compile_data(data = fits,
                     output_type = "dataframe",
                     list_element = 1)

#Compile the water mass - water potential graphs
graphs1 <- compile_data(data = fits,
                     output_type = "list",
                     list_element = 2)

#Compile the PV graphs
graphs2 <- compile_data(data = fits,
                     output_type = "list",
                     list_element = 3)
```

#8. Fitting hydraulic vulnerability curves

Current approach fits a sigmoidal model and calculates hydraulic parameters
from the curve fit. See Pammenter & Van der Willigen, 1998 and Ogle et al. 2009.
```{r}
#Read in data
data <- read.csv(system.file("extdata", "hydraulic_vulnerability.csv", 
                             package = "photosynthesis"))

#Fit hydraulic vulnerability curve
fit <- fit_hydra_vuln_curve(data[data$Tree == 5 & data$Plot == "Irrigation",],
                            varnames = list(psi = "P",
                                            PLC = "PLC"),
                            start_weibull = list(a = 2, b = 1),
                            title = "Irrigation 5")

#Return Sigmoidal model summary
summary(fit[[1]]) 

#Return Weibull model summary
summary(fit[[4]]) #expecting a = 4.99, b = 3.22

#Return model parameters with 95% confidence intervals
fit[[2]] 

#Return hydraulic parameters
fit[[3]] 

#Return graph
#fit[[5]] 

data <- unite(data, col = "ID", c("Plot", "Tree"), sep = "_")
#fit many function check to make sure it works for weibull
#Fit many curves
fits <- fit_many(data = data,
                  varnames = list(psi = "P",
                                            PLC = "PLC"),
                 group = "ID",
                 start_weibull = list(a = 4, b = 2),
                  #group = "Tree",
                  funct = fit_hydra_vuln_curve)

#To select individuals from the many fits
#Return model summary
summary(fits[[1]][[1]]) #Returns model summary

#Return sigmoidal model output
fits[[1]][[2]] 

#Return hydraulic parameters
fits[[1]][[3]] 

#Return graph
fits[[1]][[5]] 

#Compile parameter outputs
pars <- compile_data(data = fits,
                     output_type = "dataframe",
                     list_element = 3)

#Compile graphs
graphs <- compile_data(data = fits,
                     output_type = "list",
                     list_element = 5)

```

#9. Sensitivity analyses

This segment outlines a set of functions that can be used to assess the
sensitivity of data outputs to assumed parameters. For example, assuming
different values of GammaStar, mesophyll conductance, and light absorbance
on fitted gas exchange parameters.

Uncomment the chunk below to run - it takes awhile.
#```{r}
#Read in your data
#Note that this data is coming from data supplied by the package
#hence the complicated argument in read.csv()
#This dataset is a CO2 by light response curve for a single sunflower
data <- read.csv(system.file("extdata", "A_Ci_Q_data_1.csv", 
                             package = "photosynthesis"))

#Define a grouping factor based on light intensity to split the ACi
#curves
data$Q_2 <- as.factor((round(data$Qin, digits = 0)))

#Convert data temperature to K
data$T_leaf <- data$Tleaf + 273.15

#Run a sensitivity analysis on GammaStar and mesophyll conductance
#at 25 Celsius for one individual curve
pars <- analyze_sensitivity(data = data[data$Q_2 == 1500, ],
                                  funct = fit_aci_response,
                                   varnames = list(A_net = "A",
                                      T_leaf = "T_leaf",
                                      C_i = "Ci",
                                      PPFD = "Qin"),
                                   useg_mct = TRUE,
                                   test1 = "gamma_star25",
                                   element_out = 1,
                                   test2 = "g_mc25",
                                   fitTPU = TRUE,
                                   Ea_gamma_star = 0,
                                   Ea_g_mc = 0,
                                   values1 = seq(from = 20,
                                               to = 60,
                                               by = 4),
                                   values2 = seq(from = 0.2,
                                               to = 2,
                                               by = 0.1))

#Compute measures of sensitivity
par2 <- compute_sensitivity(data = pars,
                            varnames = list(Par = "V_cmax",
                                            test1 = "gamma_star25",
                                            test2 = "g_mc25"),
                            test1_ref = 40,
                            test2_ref = 1)
#Plot sensitivity
ggplot(par2, aes(y = CE_gamma_star25, x = CE_g_mc25, colour = V_cmax))+
  labs(x = expression(g_mc[25]~"Control Coefficient"),
       y = expression(Gamma[25]~"Control Coefficient")) +
  geom_point() + 
  theme_bw() 

ggplot(par2, aes(y = CE_gamma_star25, x = V_cmax, colour = gamma_star25))+
  geom_point() + 
  theme_bw() 
#Note that in this case a missing point appears due to an infinity
#Can also plot sensitivity measures in 2 dimensions
ggplot(par2, aes(x = gamma_star25, y = g_mc25, z = CE_gamma_star25))+
geom_tile(aes(fill = CE_gamma_star25)) +
labs(x = expression(Gamma*"*"[25]~"("*mu*mol~mol^{-1}*")"),
     y = expression(g[m][25]~"("*mu*mol~m^{-2}~s^{-1}~Pa^{-1}*")"))+
scale_fill_distiller(palette = "Greys") +
geom_contour(colour = "Black", size = 1) +
theme_bw()

plot(PE_gamma_star25 ~ gamma_star25, par2)
#```

#10. Dependency checking

A function can be used to generate an html file that assesses
the dependencies within and between packages
```{r}
#check_dependencies()
```


#References

Arrhenius S. 1915. Quantitative laws in biological chemistry. Bell.

Ball JT, Woodrow IE, Berry JA. 1987. A model predicting stomatal
conductance and its contribution to the control of photosynthesis
under different environmental conditions, in Progress in 
Photosynthesis Research, Proceedings of the VII International 
Congress on Photosynthesis, vol. 4, edited by I. Biggins, pp. 
221–224, Martinus Nijhoff, Dordrecht, Netherlands.

Duursma R. 2015. Plantecophys - an R package for analysing and
modeling leaf gas exchange data. PLoS ONE 10:e0143346

Erhardt EB. 2019. RLicor: Read Licor files. R package version 0.0.01.

Gu L, Pallardy SG, Tu K, Law BE, Wullschleger SD. 2010. Reliable
estimation of biochemical parameters from C3 leaf photosynthesis-
intercellular carbon dioxide response curves. Plant Cell Environ
33:1582-1874.

Harley PC, Loreto F, Di Marco G, Sharkey TD. 1992. Theoretical
considerations when estimating mesophyll conductance to CO2 flux
by analysis of the response of photosynthesis to CO2. Plant Physiol
98:1429 - 1436.

Heskel MA, O'Sullivan OS, Reich PB, Tjoelker MG, Weerasinghe LK,
Penillard A, Egerton JJG, Creek D, Bloomfield KJ, Xiang J, Sinca F,
Stangl ZR, la Torre AM, Griffin KL, Huntingford C, Hurry V, Meir P,
Turnbull MH, Atkin OK. 2016. Convergence in the temperature response
of leaf respiration across biomes and plant functional types. PNAS
113:3832-3837

Hobbs JK, Jiao W, Easter AD, Parker EJ, Schipper LA, Arcus VL.
2013. Change in heat capacity for enzyme catalysis determines
temperature dependence of enzyme catalyzed rates. ACS Chemical
Biology 8:2388-2393.

Koide RT, Robichaux RH, Morse SR, Smith CM. 2000. Plant water status, 
hydraulic resistance and capacitance. In: Plant Physiological Ecology: 
Field Methods and Instrumentation (eds RW Pearcy, JR Ehleringer, HA
Mooney, PW Rundel), pp. 161-183. Kluwer, Dordrecht, the Netherlands

Kok B. 1956. On the inhibition of photosynthesis by intense light.
Biochimica et Biophysica Acta 21: 234–244

Kruse J, Adams MA. 2008. Three parameters comprehensively describe
the temperature response of respiratory oxygen reduction. Plant 
Cell Environ 31:954-967

Laisk A. 1977. Kinetics of photosynthesis and photorespiration in
C3 plants. Nauka, Moscow.

Leuning R. 1995. A critical appraisal of a coupled stomatal-
photosynthesis model for C3 plants. Plant Cell Environ 18:339-357

Liang LL, Arcus VL, Heskel MA, O'Sullivan OS, Weerasinghe LK,
Creek D, Egerton JJG, Tjoelker MG, Atkin OK, Schipper LA. 2018.
Macromolecular rate theory (MMRT) provides a thermodynamics
rationale to underpin the convergent temperature response in
plant leaf respiration. Glob Chang Biol 24:1538-1547

Marshall B, Biscoe P. 1980. A model for C3 leaves describing the
dependence of net photosynthesis on irradiance. J Ex Bot 31:29-39

Medlyn BE, Dreyer E, Ellsworth D, Forstreuter M, Harley PC,
Kirschbaum MUF, Le Roux X, Montpied P, Strassemeyer J, Walcroft A,
Wang K, Loutstau D. 2002. Temperature response of parameters of a
biochemically based model of photosynthesis. II. A review of
experimental data. Plant Cell Environ 25:1167-1179

Medlyn BE, Duursma RA, Eamus D, Ellsworth DS, Prentice IC, Barton
CVM, Crous KY, Angelis PD, Freeman M, Wingate L. 2011. Reconciling
the optimal and empirical approaches to modeling stomatal 
conductance. Glob Chang Biol 17:2134-2144

Ogle K, Barber JJ, Willson C, Thompson B. 2009. Hierarchical statistical 
modeling of xylem vulnerability to cavitation. New Phytologist 182:541-554

Pammenter NW, Van der Willigen CV. 1998. A mathematical and statistical 
analysis of the curves illustrating vulnerability of xylem to cavitation. 
Tree Physiology 18:589-593

Sack L, Cowan PD, Jaikumar N, Holbrook NM. 2003. The 'hydrology' of 
leaves: co-ordination of structure and function in temperate woody 
species. Plant, Cell and Environment, 26, 1343-1356

Sharkey TD. 2019. Is triose phosphate utilization important for understanding
photosynthesis? Journal of Experimental Botany 70, 5521-5525

Tyree MT, Hammel HT. 1972. Measurement of turgor pressure and water 
relations of plants by pressure bomb technique. Journal of Experimental 
Botany, 23, 267

Walker BJ, Ort DR. 2015. Improved method for measuring the apparent
CO2 photocompensation point resolves the impact of multiple internal
conductances to CO2 to net gas exchange. Plant Cell Environ 38:2462-
2474

Yin X, Struik PC, Romero P, Harbinson J, Evers JB, van der Putten 
PEL, Vos J. 2009. Using combined measurements of gas exchange and
chlorophyll fluorescence to estimate parameters of a biochemical C3
photosynthesis model: a critical appraisal and a new integrated 
approach applied to leaves in a wheat (Triticum aestivum) canopy. 
Plant Cell Environ 32:448-464

Yin X, Sun Z, Struik PC, Gu J. 2011. Evaluating a new method to 
estimate the rate of leaf respiration in the light by analysis of
combined gas exchange and chlorophyll fluorescence measurements. 
Journal of Experimental Botany 62: 3489–3499